---
title: "Data Types in R "
author: Nitendra Gautam
date: November 23, 2017
output: 
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Data Types

There are five data types in R .R is case sensitive when defining variables. A variable named 'a' is not same as 'A'.


* Numeric
* Integer
* Character
* Factor
* Logical 


### Numeric Data Type 
 It includes the numbes that contains decimal point which is default .


```{r}
# Assign a value of 3 to  variable a
a <- 3

print(a)
# Check Data Types in R 
paste(" Class of A is ", class(a))

```


### Convert numeric data type into integer

```{r}
b <- as.integer(a)
paste("Class of B is " , class(b ))
```

### Character Data Type

```{r}
d <-"Nitendra Gautam"
print(d)
paste("Class of d is ",class(d))

e <- '12345'
print(e)
paste(" class of e is ",class(e))
```

 

### Factors Data Type

Factors data type refer the qualitative variables .An example can be "Good"" and "Bad".


```{r}
factorVariable <- factor(c("male","female"))

print(factorVariable)

#Factors Levels
levels(factorVariable)

#Class of the factor
class(factorVariable)

#Class of Factor Levels
class(levels(factorVariable))

#Number of Levels
nlevels(factorVariable)

```


### Logical data Type


```{r}
k <- "TRUE"
class(k)

```

### Create a factor variable with 3 Levels and 6 Items

```{r}
sV <- factor(c("good","bad","ugly","good","bad","ugly"))

class(sV)
```
