---
title: "Data Types and Variables in R "
author: Nitendra Gautam
date: November 23, 2017
output: 
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Variables and Data types in R

Everything in R is an object .
### Variable Types
* character - Strings
* integer- Integers
* numeric - Integers and Fractions
* logical -Boolean
* complex - Complex numbers 
* factor -Categorical variable where each label is a category

### Data Structures in R 

 R intuitively decides variable type in the background and assigns a class to the variable.
R has variety of data types and  is case sensitive when defining variables. A variable named 'a' is not same as 'A'.

* Atomic Vectors(numerical ,Integer,character ,logical)
* Matrix
* Data Frames
* Lists
* Factors

Here Atomic means 

R provides many functions to examine features of vectors and other object .

* ```class()``` : Gives information about the object on high level
* ```typeof()```: Gives information about the object's data type on low level
* ```length()```: Gives the length of the object
* ```attributes()```: gives the information about the metadata if the object has


## Variable in R

Variables in R are typically assigned using <- but can alos be assigned using <- or -> ,as in x <-1 or 1 -> x


```{r}

#Create variables in R and assign some value
x <- 139 +121
x

# Overwrite variables in R
x <- x/60
x

# Assign Character Values
myName <- "Nitendra Gautam"

"Nitendra Gautam" -> newName

typeof(myName) # Gives the data type of myName

# To remove the variable from Memory use rm(variable_name) command
# Removes the variables x from Memory 
rm(x)
```


## Order of Operation in R 

In order to put emphasis on order of operations use the small bracket.

For example : ```totalHours <- (139 + 121) /60```


### Numeric Data Type 
 It includes the numbes that contains decimal point which is default .


```{r}
# Assign a value of 3 to  variable a
a <- 3

print(a)
# Check Data Types in R 
paste(" Class of A is ", class(a))

# Convert numeric data type into integer
b <- as.integer(a)
paste("Class of B is " , class(b ))
```



### Character Data Type

```{r}
d <-"Nitendra Gautam"
print(d)
paste("Class of d is ",class(d))

e <- '12345'
print(e)
paste(" class of e is ",class(e))
```

 

### Factors Data Type

Factors data type refer the qualitative variables .An example can be "Good"" and "Bad".


```{r}
factorVariable <- factor(c("male","female"))

print(factorVariable)

#Factors Levels
levels(factorVariable)

#Class of the factor
class(factorVariable)

#Class of Factor Levels
class(levels(factorVariable))

#Number of Levels
nlevels(factorVariable)

```

### Logical data Type

```{r}
k <- "TRUE"
class(k)

# Create a factor variable with 3 Levels and 6 Items
sV <- factor(c("good","bad","ugly","good","bad","ugly"))

class(sV)

levels(sV)
```


### Vectors

Vector are type of object which is used to store multiple data of same data type

Technically, vectors can be one of two types:

* atomic vectors
* lists

In general vector refers to the atomic types not the lists 
Vector can have character ,integer,numeric or logical value.


```{r}

v1 <- c(1,2,5.3,6,-2,4,5) # create a Numeric Vector 
v2 <- c("one","two","three","four","five") # Character Vector
v3 <- c(TRUE,FALSE,TRUE,FALSE) # Logical Vector

# Refer to the element of the vector using Subscript 
v1[c(3,4)]  #Third and Fourth Elements of Vector

v1[4] # Get Fourth Element from Vector V1

v1[2:6]  # Get elements fomr 2nd to 6th element 

class(v2) # Class of Character Vector

v2[2:5] # Get elements from second to 5th element

 typeof(v2)  #Gives the types of Vector V2
 
```

### Adding Elements in character Vector

```{r}

z <-c("Sam","Ram","Hari")

## Add oneelements in Vector
z <- c(z,"Mike")

## Add Another elements in Vector
z <- c(z,"Shyam")
```


### Given a Large Vector find the first and Last Element of the Vector

```{r}
ab <- c(1:50) # Large vector from 1 to 50

ab[1] # Get the first Element 
ab[length(ab)] # Get the last element which will be at the index equal to length of Vector

ab[c(ab[1],ab[length(ab)])]
```



## References

[1]  <a href="https://swcarpentry.github.io/r-novice-inflammation/13-supp-data-structures/">R Data Structures </a>



