---
title: "Data Types and Variables in R "
author: Nitendra Gautam
date: November 23, 2017
output: 
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Variable in R

Variables in R are typically assigned using <- but can alos be assigned using = ,as in x <-1 or x =1


* Create variables in R and assign some value

```{r}
x <- 139 +121
x
```


* Overwrite variables in R






## R Data Types



R has variety of data types and  is case sensitive when defining variables. A variable named 'a' is not same as 'A'.

* Scalar
* Vectors(numerical ,Integer,character ,logical)
* Matrices
* Data Frames
* Lists
* Factor
* Logical 


### Numeric Data Type 
 It includes the numbes that contains decimal point which is default .


```{r}
# Assign a value of 3 to  variable a
a <- 3

print(a)
# Check Data Types in R 
paste(" Class of A is ", class(a))

```


### Convert numeric data type into integer

```{r}
b <- as.integer(a)
paste("Class of B is " , class(b ))
```

### Character Data Type

```{r}
d <-"Nitendra Gautam"
print(d)
paste("Class of d is ",class(d))

e <- '12345'
print(e)
paste(" class of e is ",class(e))
```

 

### Factors Data Type

Factors data type refer the qualitative variables .An example can be "Good"" and "Bad".


```{r}
factorVariable <- factor(c("male","female"))

print(factorVariable)

#Factors Levels
levels(factorVariable)

#Class of the factor
class(factorVariable)

#Class of Factor Levels
class(levels(factorVariable))

#Number of Levels
nlevels(factorVariable)

```


### Logical data Type


```{r}
k <- "TRUE"
class(k)

```

### Create a factor variable with 3 Levels and 6 Items

```{r}
sV <- factor(c("good","bad","ugly","good","bad","ugly"))

class(sV)

levels(sV)
```


### Vectors

Vector are type of object which is used to store multiple data of same data type


```{r}

v1 <- c(1,2,5.3,6,-2,4,5) # Numeric Vector 
v2 <- c("one","two","three","four","five") # Character Vector
v3 <- c(TRUE,FALSE,TRUE,FALSE) # Logical Vector
```

* Refer to the element of the vector using Subscript 

```{r}
v1[c(3,4)]  #Third and Fourth Elements of Vector

v1[4] # Get Fourth Element from Vector V1

v1[2:6]  # Get elements fomr 2nd to 6th element 

class(v2) # Class of Character Vector

v2[2:5] # Get elements from 
```


### Given a Large Vector find the first and Last Element of the Vector

```{r}
ab <- c(1:50) # Large vector from 1 to 50

ab[1] # Get the first Element 
ab[length(ab)] # Get the last element which will be at the index equal to length of Vector

ab[c(ab[1],ab[length(ab)])]
```





