---
title: "Basic Math with R"
author: "Nitendra Gautam"
date: November 23 ,2017
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R have following types of Operators 

* Arithmetic Operators
* Relational Operators
* Logical Operators
* Assignment Operators
* Miscellaneous Operators



## Arithmetic Operators 

* Square root

```{r}
sqrt(25)
```


* Natural Logarithm 

```{r}
log(100)
```

* Log of base 10

```{r}
log10(10)
```

* Power of Number
```{r}
3 ^3
```

```{r}
3^9
```


* Value of pi

```{r}
pi
```



* Rounding a number
```{r}
round(pi * 3)
```

* Rounding up number to upper bound /nearest higher Integer

```{r}
ceiling(pi * 3)
```

* Rounding to the nearest lower Integer

```{r}
floor(pi * 3)
```


* Finding Quotient of specific division 

```{r}
25 %/% 6
```

* Finding remainder of two numbers 
```{r}
125 %% 6
```


* Square root of 729
```{r}
sqrt(729)
```


* Remainder of 109/25

```{r}
109 %% 25
```

